---
title: "<center> <h1>STT 3850 Midterm Study Guide</h1> <center>"
author: "Andrew Thorp"
date: "October 9, 2017"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

#Statistics

###Characterizing a set of data (3 S’s)
  * Shape: how the data is distributed
  * Low outliers make a dataset skewed to the Right
  * High outliers make a dataset skewed to the Left
  * Normal distributions have fairly even outliers on either side
  
###Center: Where the data is centered around
  * Normal: If the dataset has a normal distribution (shape) this can be calculated using the `mean($data)` function.
  * Skewed (left or right): The mean will be misrepresent the center. Calculate a skewed center using `median($data)`.

###Spread: How far the data differs from the center
  * Normal: If the dataset has a normal distribution, then the standard deviation applies to both sides of the data and so it represents the spread.
  * Skewed (left or right): If the dataset is skewed on either side, the deviation above and below the center will not be the same, so you must calculate it using `IQR($data)` for the interquartile range.

###Hypothesis testing (5 step procedure)
  * Z-Score: The number of Standard deviations an element is form the mean.
  * P-Score: 
  * $\bar{X}=mean$
  
  1. Specify the Null and ALternative hypothesis
    + Null hypothesis notated as $H_0:M=value$ or $\bar{M_1}-\bar{M_2}=0$
    + Alternative hypothesis notated as $H_A:M\neq0$
    
  2. Test your staitstic using the Z-test or T-test
    + `t.test(variable~catagoricalVariable, data=DF)` will perform the t-test on a set of data. If the data is not tidy your might want to use `dplyr` to tidy it up first. 
    
    Example:
```{r}
DF <- ChickWeight %>%                          #imports data frame
  filter(Diet %in% c(3:4))                     #then removes all collums except for 3 and 4
                                               #weight~Diet means the weight value grouped by the Diet value.
t.test(weight~Diet, data=DF)                   # Quickly gives us what we need to know
```




Markdown
Dplyr
Ggplot2
