---
title: "<center> <h1>STT 3850 Midterm Study Guide</h1> <center>"
author: "Andrew Thorp"
date: "October 9, 2017"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(lsr)
```

#Statistics

###Characterizing a set of data (3 S’s)
######Shape: how the data is distributed
  * Low outliers make a dataset skewed to the Right
  * High outliers make a dataset skewed to the Left
  * Normal distributions have fairly even outliers on either side
  
######Center: Where the data is centered around
  * Normal: If the dataset has a normal distribution (shape) this can be calculated using the `mean($data)` function.
  * Skewed (left or right): The mean will be misrepresent the center. Calculate a skewed center using `median($data)`.

######Spread: How far the data differs from the center
  * Normal: If the dataset has a normal distribution, then the standard deviation applies to both sides of the data and so it represents the spread.
  * Skewed (left or right): If the dataset is skewed on either side, the deviation above and below the center will not be the same, so you must calculate it using `IQR($data)` for the interquartile range.

###Hypothesis testing (5 step procedure)
  * Z-Score: The number of Standard deviations an element is form the mean.
  * P-Score: 
  * $\bar{X}=mean$
  
  1. Specify the Null and ALternative hypothesis
    + Null hypothesis notated as $H_0:M=value$ or $\bar{M_1}-\bar{M_2}=0$
    + Alternative hypothesis notated as $H_A:M\neq0$
    
  2. Test your staitstic using the Z-test or T-test
    + `t.test(variable~catagoricalVariable, data=DF)` will perform the t-test on a set of data. If the data is not tidy your might want to use `dplyr` to tidy it up first. 
    + Alternatively, if we have an expected mean, and our data has the experimental mean, we can call the `t.test` function as such: `t.test(experimental_data, mu = <expected_mean>)`
    Example:
```{r, results='hide'}
DF <- ChickWeight %>%                 # Imports data frame
  filter(Diet %in% c(3:4))            # Then removes all collums except for 3 and 4
                                      # Weight~Diet means the weight across Diets value.
t.test(weight~Diet, data=DF)          # Quickly gives us what we need to know
```
    
  3. Determine rejection region
    * Rejection region is the percent of data points on either end of a data set.
    * Percentile is represented by alpha ($\alpha$)
    * Unless otherwise specified, assume rejection region to be 5% (0.05).
    
    
  4. Statistical conclusion. 
     
     You can either
      * Reject the null hypothesis $H_0$ if $P<\alpha$
      
      or
      * Fail to reject the null hypothesis $H_0$ if $P\geq\alpha$
      
      Based on the output from the previous example pictured below, P = 0.45, and we assume $\alpha$=0.05. P is far greater than $\alpha$, so we **fail to reject the null hypothesis** $H_0$.
      
```{r, echo=FALSE}
DF <- ChickWeight %>%                 # Imports data frame
  filter(Diet %in% c(3:4))            # Then removes all collums except for 3 and 4
                                      # Weight~Diet means the weight across Diets value.
t.test(weight~Diet, data=DF)          # Quickly gives us what we need to know
```

      
  5. English conclusion (or Spanish, if that's your thing)
    * The statistical conclusion deals with the null hypothesis $H_0$, but the English conclusion only deals with things in terms of the Alternative Hypothesis $H_A$ (written out)
    
    In the case of our example from above, we failed to reject the null hypothesis, so we **failed to find evidence supporting a weight difference in chicks between Diets 3 and 4**.
    * The English hypothesis must ALWAYS be written this way. You either find evidence to support $H_A$ or you fail to find evidence to support $H_A$.

### Cohens'd (effect size)
  * The difference between a control group's mean, and an experiment group's mean, measured in standard deviations. 
  * Measured using the following formula: $d=\frac{|M_0 - M_1 |}{SD}$
  * How large is the effect size?
    + small if    |d| < 0.20
    + medium if   |d| < 0.5
    + large if    |d| > 0.5
  * Compute using R with `cohensD(post_data, pre_data, method = "paired")`
```{r, eval = FALSE}
Diet_3 <- filter(ChickWeight, Diet == 3)
Diet_4 <- filter(ChickWeight, Diet == 4)
#method should normally be "paired", but the sample size was different
cohensD(Diet_4$weight, Diet_3$weight, method = "unequal") 
```
  
### Analysing a DataSet
  * An analysis consists of 
    + t-value
    + p-value
    + effect size (Cohen's D)
```{r, eval = FALSE}
DF <- ChickWeight %>% 
  filter(Diet %in% c(3:4))
t.test(weight ~ Diet, data = DF)

Diet_3 <- filter(DF, Diet == 3)
Diet4 <- filter(DF, Diet == 4)
  cohensD(Diet_3$weight, Diet_4$weight, method = "unequal")


```
    From this we can gather the effect size is roughly 10% and, with a t-value of 75% and a p-value of 45%.
    TODO: ask about reading these statistics.
    
### 

Dplyr
Ggplot2
